---
title: "A Report on Lake Ontario's Microbes"
author: "Rachel Whiteside"
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction

In this report, we will analyze how different water quality parameters affect the abundance of bacterial phyla.

# Set-up

## Loading packages

First, we'll load necessary libraries.

```{r packages, message=FALSE}
library(tidyverse)
```

## Reading in data

Then, we'll read in our data. We produced this dataframe by joining the sample_data.csv and taxon_abundance.csv files using innerjoin.

```{r data, message=FALSE}
sample_and_taxon <- read_csv('data/sample_and_taxon.csv')
```

## Analyzing the data

Finally, we'll plot the relationship between temperature and overall microbial cell abundance.

```{r cell_vs_temp}
sample_and_taxon %>% 
  ggplot() +
  aes(x = temperature, 
      y = cells_per_ml/1e6, 
      color = env_group) + 
  geom_point() + 
  labs(x = 'Temperature (C)',
       y = 'Cells (million/ml)',
       title =  'Are temperature and cell abaundance linked?')
```

It appears that warmer waters promote higher microbial cell counts. Cell counts were especially high in the warm waters of September.

```{r data_summary, include=FALSE}
nSamples <- nrow(sample_and_taxon)

minTemp <- sample_and_taxon %>%
  pull(temperature) %>%
  min() %>%
  round()

#make an object called maxTemp which has maximum temp

maxTemp <- sample_and_taxon %>%
  pull(temperature) %>%
  max() %>%
  round()
```

The above plot shows `r nSamples`. For these samples, the minimum temp was `r minTemp`C and the maximum temp was `r maxTemp`C.

| Statistic         | Value        |
|-------------------|--------------|
| Number of Samples | `r nSamples` |
| Minimum Temp      | `r minTemp`  |
| Maximum Temp      | `r maxTemp`  |

```{r, stargazer, echo=FALSE}
install.packages("stargazer")
library(stargazer)

```

```{r kable}
library(knitr)

sample_and_taxon %>% 
  filter(env_group == 'Deep') %>% 
  select(cells_per_ml, temperature, Chloroflexi) %>% 
  kable()
```

Challenge: Make a bullet point list First line is your favorite coding language - italics Second line is favorite function - bold Third line is the next programming thing you want to learn - italics and bold

-   *Favorite coding language is R*

-   **Favorite function is summarize**

-   ***Next thing I want to learn is Geospatial Analysis in R***

1.  *Favorite coding language is R*

2.  **Favorite function is summarize**

3.  ***Next thing I want to learn is Geospatial Analysis in R***

then, switch to numbered then, find markdown formatting tutorial and insert as a hyperlink

[Here is a link to table formatting options](https://rmarkdown.rstudio.com/lesson-7.html)
